Customizing Software by Direct Manipulation of Tabular Data

### 2020-05-22 paper.tex
actually finished 3.1

\textbf{Adding notes to links}: The user can also manually add notes to
the table, by simply entering values into the table without using
formulas. In this case, the user jots down a few notes in another column
about articles they might want to read, and the notes appear in the page
next to the read times (Figure~\ref{fig:hacker-news}, Note D).

@@

Figure~\ref{fig:table-adapter} summarizes the overall architecture of
table-driven customization, using a simplified version of the Airbnb
example above. (\emph{todo: change this figure to hacker news}) In this
example, the name of each listing is scraped from the web page DOM, the
latitude and longitude of each listing is scraped from AJAX responses,
and user annotations are loaded from the brower's local storage.

First, the three different data sources are each bidirectionally mapped
to a table interface by a \textbf{table adapter}. The table adapter
defines how to map a particular type of data to a table, and what
effects edits should have on the original data source. In some cases,
the mapping logic is straightforward: the local storage adapter stores a
table of data, so the mapping to the table abstraction is trivial. In
other cases, the mapping is more involved: the DOM scraping adapter
implements web scraping logic to produce a table of data from the web
page, and turns edits to the table into DOM manipulations like
reordering rows of data on the page.

The three separate tables are then combined into a single table for the
end user to view and edit. The \textbf{query engine} is responsible for
creating this combined view, and routing the user's edits back to the
individual table adapters. In this example, the query engine has joined
the three tables together by a shared ID column, and sorted the result
by the name column.

@@

A key idea in table-driven customization is that a wide variety of data
sources can be mapped to the generic abstraction of an editable table.
In a relational database, the table matches the underlying storage
format, but in table-driven customization, the table is merely an
\emph{interface layer}. The data shown in the table is a projection of
some underlying state, and edits to the table can have complex effects
on the underlying state.

Externally, a table adapter must satisfy an abstract interface. The
first two parts of the interface resemble the interface exposed by a
table in a relational database:

\textbf{Returning a table}: A table adapter exposes a table of data: an
ordered list of records. Each record carries a unique identifier and
associates named attributes with values. Tables have a typed schema, so
the same attributes are shared across all records. A table adapter can
update the contents of a table at any time, in response to changes in
the underlying state (e.g., a DOM scraping adapter can update the table
when the page body changes). When data changes, the query view is
reactively updated in response.

\textbf{Making edits}: The query engine can issue a request to a table
adapter to make an edit to a record. The meaning of making an edit can
vary depending on the adapter: in the local storage adapter, an
annotation may be persisted into local storage; in the DOM scraping
adapter, an edit may represent filling in a form field. An adapter can
also mark values as read-only if it wouldn't be meaningful to edit them;
for example, the DOM scraping adapter typically marks page content as
read-only, except for editable form fields.

The query engine also sends additional information about the combined
query view to each table adapter. These functions are currently used to
provide the DOM scraping adapter with sufficient information to
manipulate the original application UI as the user manipulates the table
view.

\textbf{Sorting/filtering}: When the user sorts or filters the query
view, an ordered list of visible IDs is sent to each table adapter. The
DOM scraping adapter uses this information to change the list of rows
shown in the web page.

\textbf{Information from other tables}: The query engine sends each
table adapter the entire combined table of data being rendered to the
user. The DOM scraping adapter uses this to add annotations to the page
by looking for additional data columns joined to scraped rows, and
rendering them in the page. (\emph{todo: explain joined})

\textbf{Currently selected record}: The query engine notifies each table
adapter about which record is currently selected by the user in the
table UI. The DOM scraping adapter uses this information to highlight
the row in the page that corresponds to the selected record in the
table. (\emph{todo: clarify})

@@

Here we describe in more detail the table adapters that we have
implemented in Wildcard to power the customizations shown in
Section~\ref{sec:example}.

@@

Because each website has unique content, we rely on programmers to
create a DOM scraping adapter for each individual website to make it
available for customization in Wildcard. To make this approach viable,
we have built a library of functions that make it easy to implement a
scraping adapter, requiring the programmer only to implement the minimal
site-specific parts. The programmer writes a single Javascript function
which can use standard scraping techniques like CSS selectors and DOM
APIs to extract relevant elements from the page. The library functions
then wrap this function to implement the rest of the needed
functionality; for example, when the table is sorted, we find the DOM
elements corresponding to the rows in the table, remove them from the
page, and then reinsert them in the new order.

An \textbf{AJAX scraping adapter} intercepts AJAX requests made by a web
page, and extracts information from those requests. When available, this
tends to be a helpful technique because the data is already in a
structured form, and often includes information not shown in the UI. As
with DOM scraping adapters, we have made it easy for programmers to
create site-specific AJAX scraping adapters. A programmer writes a
function that specifies how to extract data from an AJAX request, and
the framework handles the details of intercepting requests and calling
the programmer-defined function.\footnote{So far we have only
  implemented AJAX scraping in the Firefox version of Wildcard, since
  Firefox has convenient APIs for intercepting requests. It appears
  possible to implement in Chrome as well, but we have not finished our
  implementation.}

The \textbf{local storage adapter} simply stores a table of data in the
browser. As shown in the example use cases, this can be useful for
persisting annotations in a private way, without uploading them to a web
service.

@@

\emph{Exposing a table}: In Redux, app developers already represent the
state of a user interface as a single object. To expose a table of data,
the developer simply writes a function that maps the state object to a
table. \emph{Editing data values}: The developer defines how edits to
records in the table should map to edits in the state object.
\emph{Sorting/filtering the UI}: The developer inserts an additional
data processing step before the state object is rendered in the UI, so
that Wildcard can apply sorting and filtering. \emph{Annotations}: The
developer adds additional React components managed by Wildcard into
their UI tree, which render additional data as annotations.

@@

>>>>

\textbf{Adding notes to links}: The user can also manually add notes to
the table, by simply entering values into the table without using
formulas. In this case, the user jots down a few notes in another column
about articles they might want to read, and the notes appear in the page
next to the read times (Figure~\ref{fig:hacker-news}, Note D). The
annotations are also stored in the browser's local storage so they can
be retrieved on future visits.

@@

Figure~\ref{fig:table-adapter} summarizes the overall architecture of
table-driven customization, using a simplified illustration of the
Hacker News example scenario. In this example, the name and points value
for each article is scraped from the web page DOM, and user annotations
are loaded from the brower's local storage.

First, the web page and the browser storage are each wrapped by a
\textbf{table adapter}, which defines a bidirectional mapping between
some underlying data source and a table.

In addition to a \emph{read} mapping for how the underlying data should
be represented as a table, it also defines a \emph{write} mapping for
what effects edits should have on the original data source. In this
case, the local storage adapter has a trivial mapping: it loads a table
of data stored in the browser, and persists edits to that persisted
state. The mapping logic of the DOM scraping adapter is much more
involved. It implements web scraping logic to produce a table of data
from the web page, and turns edits into DOM manipulations, such as
reordering rows of data on the page.

The two tables are then combined into a single table for the end user to
view and edit. The \textbf{query engine} is responsible for creating
this combined view, and routing the user's edits back to the individual
table adapters. In this example, the query engine has joined the two
tables together by a shared ID column, and sorted the result by the
points column.

@@

A key idea in table-driven customization is that a wide variety of data
sources can be mapped to a generic table abstraction. In a relational
database, the table matches the underlying storage format, but in
table-driven customization, the table is merely an \emph{interface
layer}. The data shown in the table is a projection of some underlying
state, and edits to the table can have complex effects on the underlying
state.

The first two parts of the table adapter interface resemble a typical
database table:

\emph{Returning a table}: A table adapter exposes a table of data: an
ordered list of records. Each record carries a unique identifier and
associates named attributes with values. Tables have a typed schema, so
the same attributes are shared across all records. The columns also
carry some additional metadata in addition to their type, such as
whether or not they are read-only or editable.

A table adapter can update the contents of a table at any time in
response to changes in the underlying state (e.g., a DOM scraping
adapter can update the table when the page body changes). When data
changes, the query view is reactively updated in response.

\emph{Making edits}: The query engine can issue a request to a table
adapter to make an edit to a record. The meaning of making an edit can
vary depending on the adapter: in the local storage adapter, an
annotation may be persisted into local storage; in the DOM scraping
adapter, an edit may represent filling in a form field.

In addition, the query engine also sends additional information about
the combined query view to each table adapter:

\emph{Sorting/filtering}: When the user sorts or filters the query view,
an ordered list of visible IDs is sent to each table adapter. The DOM
scraping adapter uses this information to change the list of rows shown
in the web page.

\emph{Data from other tables}: The query engine provides each table
adapter with the entire combined table shown to the user. The DOM
scraping adapter uses this for injecting annotations---values in
additional columns from other tables are added to the original web page.

\emph{Currently selected record}: As the user clicks in different parts
of the table view, the query engine communicates the record currently
selected by the user. The DOM scraping adapter uses this information to
highlight the row in the page that corresponds to the selected row in
the table (and scroll to the row if it's not currently showing on the
page), which helps clarify the connection between the table and the
original UI.

@@

In Wildcard, we have built three specific types of table adapters so
far.

@@

Because each website has unique content, we rely on programmers to
create a DOM scraping adapter for each individual website to make it
available for customization in Wildcard. To make this approach viable,
we have built a generic DOM scraping adapter which a programmer can
configure with the minimal site-specific parts.

After specifying basic metadata like which URLs the adapter should apply
to, the programmer only needs to implement a single Javascript function
that scrapes relevant elements from the page. They are free to use any
scraping techniques; in practice simple CSS selectors and DOM APIs
usually suffice. he generic adapter then wraps this scraping function to
implement the table adapter interface. For example, when the table view
is sorted, the generic adapter takes the DOM elements corresponding to
the rows in the table, removes all of them from the page, and then
reinserts them in the new order.

The programmer can optionally override other default behavior of the
generic DOM scraping adapter, including the logic for when to re-run the
scraping function in response to page changes, how to style injected
annotations, and how to style the currently selected row in the table.

An \textbf{AJAX scraping adapter} intercepts AJAX requests made by a web
page, and extracts information from those requests to add to the table.
When available, this tends to be a helpful technique because the data is
already in a structured form so it is easier to scrape, and it often
includes valuable information not shown in the UI.

As with DOM scraping adapters, we have made it easy for programmers to
create site-specific AJAX scraping adapters. A programmer writes a
function that specifies how to extract data from an AJAX request, and
the framework handles the details of intercepting requests and calling
the programmer-defined function.\footnote{So far we have only
  implemented AJAX scraping in the Firefox version of Wildcard, since
  Firefox has convenient APIs for intercepting requests. It appears
  possible to implement in Chrome as well, but we have not finished our
  implementation.}

In order to join the tables produced by AJAX scraping and DOM scraping,
a common set of identifiers is required across records in the two
tables. Often there is a server-defined ID present both in the DOM and
in AJAX responses; if not, the programmer can use some set of
overlapping data (e.g.~an item name) as a shared ID.

The \textbf{local storage adapter} simply stores a table of data in the
browser. This is currently only used to persist annotations: the user
makes edits in the table, they are stored privately in the browser, and
can be loaded into the table on future pageloads.

@@

\begin{itemize}
\tightlist
\item
  \emph{Exposing a table}: In Redux, app developers already represent
  the state of a user interface as a single object. To expose a table of
  data, the developer simply writes a function that maps the state
  object to a table.
\item
  \emph{Editing data values}: The developer defines how edits to records
  in the table should map to edits in the state object.
\item
  \emph{Sorting/filtering the UI}: The developer inserts an additional
  data processing step before the state object is rendered in the UI, so
  that Wildcard can apply sorting and filtering.
\item
  \emph{Annotations}: The developer adds additional React components
  managed by Wildcard into their UI tree, which render additional data
  as annotations.
\end{itemize}

@@

\emph{todo: mention ajax scraping} \emph{todo: mention privacy benefits}
